---
title: 'Growth Curve FittR v1.0'
output: 
  flexdashboard::flex_dashboard:
    source_code: https://github.com/saunders-lab/growth_curve_fittr
    theme: bootstrap
fontsize: 12pt
runtime: shiny
---

```{r setup}
library(tidyverse)
#library(dplyr)
#library(cowplot)
#library(viridis)
#library(knitr)
library(modelr)
library(broom)
#ibrary(ggforce)
library(grofit)


theme_set(theme_bw())
```

# Instructions 

```{r, child = 'app_embedded_instructions.Rmd'}
res <- knitr::knit_child('app_embedded_instructions.Rmd', quiet = TRUE)
cat(res, sep = '\n')
```

# 1. Upload Data { data-orientation=rows}

## {.sidebar data-width='300%'}

### 1. Upload Data

See **[Instructions]** for more details. 

Upload a .csv file where column 1 is "time" (in hrs). Data columns for each well / experiment follow. 

#### Upload csv file:

```{r}
fileInput("file1", label = NULL, multiple = FALSE, 
          accept = c("text/csv", "text/comma-separated-values,text/plain",".csv"))

df <- reactive({
  
  req(input$file1)
  
  df <- read_csv(input$file1$datapath)
  
  df_well_index <- tibble(well = colnames(df)[-1]) %>% mutate(well_index = row_number())

  df_wIndex <- left_join(df %>% gather(key = 'well',value = 'absorbance',-time), 
                         df_well_index, 
                         by = 'well' )
  
  
  return(df_wIndex)
  
})

```

Here's are some [example datasets](). Download the datasets and choose one file to upload. 

## Plots


### Overview Plot

```{r}
renderPlot({
  
  ggplot(df(), aes(time, absorbance, group = well)) + geom_path() + theme_bw()
  
})
```

## Tables

### Overview Table

```{r}
renderTable({
  
  df() %>% pivot_wider(id_cols = 'time', names_from = 'well', values_from = 'absorbance' )
  
})
```


# 2. Select Data to Fit { data-orientation=rows}

## {.sidebar data-width='300%'}

### 2. Select Data to Fit

## Plots {data-height='300%'}

### Plot Selected Curves


```{r}

output$selection_plot <- renderPlot({
  
  ggplot(df() %>% filter(well %in% input$selected_wells), aes(time, absorbance, group = well)) + geom_path() + labs(title = 'Drag the box to select data')+ theme_bw()
  
})

plotOutput("selection_plot", brush = brushOpts(id = "selection_plot_brush"))

# Reactive that returns the whole dataset if there is no brush

selectedData <- reactive({
  data <- brushedPoints(df(), input$selection_plot_brush) 
  if (nrow(data) == 0)
    data <- df()
  data %>% filter(well %in% input$selected_wells)
})

```

### Plot Selected Data

```{r}

#output$select_plot <- 
  
  renderPlot({
  
  ggplot(selectedData(), aes(time, absorbance, color = well)) + geom_path() + scale_color_viridis_d() + theme(legend.position = 'bottom')+ theme_bw()
  
})

#plotOutput("select_plot", height = '300px')

#renderUI(list(tags$hr(),plotOutput("overview_plot", brush = brushOpts(id = "overview_plot_brush"), height = '50px'), tags$hr(), plotOutput("select_plot", height = '100px')))

```


## Select Wells{data-height='100%'}

### Select Subset of Growth Curves

```{r}
renderUI({
  list(
    sliderInput(inputId = 'well_range', label = 'Select a range of curves', min = 1, max = max(df()$well_index), value = c(1,max(df()$well_index)))
  )
})

```

### Select Individual Curves

```{r}
renderUI({    checkboxGroupInput(inputId = 'selected_wells', label = 'Choose specific curves',
                     choices = levels(factor((df() %>% filter(well_index >= input$well_range[1] & well_index <= input$well_range[2]))$well)),
                     selected = levels(factor((df() %>% filter(well_index >= input$well_range[1] & well_index <= input$well_range[2]))$well)),
                     width = '100%', inline = T)})

```

## Tables {data-height='100%' .tabset}

### Table Selected


```{r}
renderTable({
  selectedData() %>% pivot_wider(id_cols = 'time', names_from = 'well', values_from = 'absorbance' )
})
```



# 3. Fit Models {data-orientation=rows}

## {.sidebar data-width='300%'}

### 3. Fit Models

See Instructions for more details.

#### Select a Model:

```{r}
selectInput("model", NULL, c("Single Gompertz" = "gomp", "Bi-Gompertz" = "biGomp","Spline" = "spline"), selected = 'gomp')
```

* **Single Gompertz:** assumes a single growth phase
* **Bi-Gompertz:** assumes two simultaneous growth phases
* **Spline:** model-free approach, assumes nothing

#### Parameters:

* **A:** the asymptote of the maximum value.
* **Mu / $\mu$:** (Maximum specific growth rate) the slope of the tangent at the inflection point.
* **Lag / $\lambda$:** (Lag time) the x intercept of the Mu tangent.
* **C:** minimal value offset from zero.

#### Set starting parameter values

```{r}

params_gomp <- list(
      numericInput("A", "A start value", 0.5), 
      numericInput("Lag", "Lag start value", 4),
      numericInput("Mu", "Mu start value", 0.2), 
      numericInput("C", "C start value", 0.05),
      NULL,
      NULL,
      NULL
    )

params_biGomp <- list(
      numericInput("A", "A_1 start value", 0.5), 
      numericInput("A_2", "A_2 start value", 0.2),
      numericInput("Lag", "Lag_1 start value", 4),
      numericInput("Lag_2", "Lag_2 start value", 5),
      numericInput("Mu", "Mu_1 start value", 0.2), 
      numericInput("Mu_2", "Mu_2 start value", 0.3),
      numericInput("C", "C start value", 0.05)
    )

params_spline <- list(renderText("No params needed"),NULL, NULL, NULL, NULL, NULL, NULL )


conditionalPanel(condition = "input.model == 'gomp'", 
                 params_gomp
                 )

conditionalPanel(condition = "input.model == 'biGomp'", 
                 params_biGomp
                 )

conditionalPanel(condition = "input.model == 'spline'", 
                 params_spline
                 )


```



#### Set parameter bounds

```{r}
bounds_gomp <- list(
  numericInput("A_low", "A Lower bound", 0),
  numericInput("A_up", "A Upper bound", 100),
  numericInput("Lag_low", "Lag Lower bound", 0),
  numericInput("Lag_up", "Lag Upper bound", 100),
  numericInput("Mu_low", "Mu Lower bound", 0),
  numericInput("Mu_up", "Mu Upper bound", 10),
  numericInput("C_low", "C Lower bound", -10),
  numericInput("C_up", "C Upper bound", 10),
  NULL,
  NULL,
  NULL,
  NULL,
  NULL,
  NULL
)

bounds_biGomp <- list(
  numericInput("A_low", "A_1 Lower bound", 0),
  numericInput("A_up", "A_1 Upper bound", 100),
  numericInput("A_2_low", "A_2 Lower Bound", 0),
  numericInput("A_2_up", "A_2 Upper Bound", 100),
  numericInput("Lag_low", "Lag_1 Lower bound", 0),
  numericInput("Lag_up", "Lag_1 Upper bound", 100),
  numericInput("Lag_2_low", "Lag_2 Lower Bound", 0),
  numericInput("Lag_2_up", "Lag_2 Upper Bound", 100),
  numericInput("Mu_low", "Mu_1 Lower bound", 0),
  numericInput("Mu_up", "Mu_1 Upper bound", 10),
  numericInput("Mu_2_low", "Mu_2 Lower Bound", 0),
  numericInput("Mu_2_up", "Mu_2 Upper Bound", 10),
  numericInput("C_low", "C Lower bound", -10),
  numericInput("C_up", "C Upper bound", 10)
)

bounds_spline <- list( 
  renderText("No bounds needed"),
  NULL,
  NULL,
  NULL,
  NULL,
  NULL,
  NULL,
  NULL,
  NULL,
  NULL,
  NULL,
  NULL,
  NULL,
  NULL
)

conditionalPanel(condition = "input.model == 'gomp'", 
                 bounds_gomp
                 )

conditionalPanel(condition = "input.model == 'biGomp'", 
                 bounds_biGomp
                 )

conditionalPanel(condition = "input.model == 'spline'", 
                 bounds_spline
                 )

```




```{r}
# This is where the models are actually fit

df_models <- reactive({
        
  
        # group and nest the data
        df_nest <- selectedData() %>%
            group_by(well) %>% 
            nest()
        
        
        # The nls function used to fit the data

        
        fit_gompertz_mod <- function(df){
            
            #A = 1; mu = 0.1; lam = 2
            
            mod <- nls(formula=absorbance~A*exp(-exp(((Mu*exp(1))/A)*(Lag - time) + 1))+C, 
                       start = c(A = input$A, Mu = input$Mu, Lag = input$Lag, C = input$C), 
                       algorithm = 'port',
                       lower = c(A = input$A_low, Mu = input$Mu_low, Lag = input$Lag_low, C = input$C_low),
                       upper = c(A = input$A_up, Mu = input$Mu_up, Lag = input$Lag_up, C = input$C_up),
                       data = df)
            
            mod
        }
        
        fit_biGompertz_mod <- function(df){
            
            mod <- nls( formula=absorbance~( A_1*exp(-exp(((Mu_1*exp(1))/A_1)*(Lag_1 - time) + 1)) ) + 
                                           ( A_2*exp(-exp(((Mu_2*exp(1))/A_2)*(Lag_2 - time) + 1)) ) + C,
                        
#                        start = c(A_1 = 0.5, Mu_1 = 0.2, Lag_1 = 4, C= 0.05, 
#                                  A_2 = 0.2, Mu_2 = 0.3, Lag_2 = 5),
                        start = c(A_1 = input$A, Mu_1 = input$Mu, Lag_1 = input$Lag, C = input$C, 
                                  A_2 = input$A_2, Mu_2 = input$Mu_2, Lag_2 = input$Lag_2),
                        lower = c(A_1 = input$A_low, A_2 = input$A_2_low, Mu_1 = input$Mu_low, Mu_2 = input$Mu_2_low, Lag_1 = input$Lag_low, Lag_2 = input$Lag_2_low, C = input$C_low),
                        upper = c(A_1 = input$A_up, A_2 = input$A_2_up, Mu_1 = input$Mu_up, Mu_2 = input$Mu_2_up, Lag_1 = input$Lag_up, Lag_2 = input$Lag_2_up, C = input$C_up),
                        data = df, 
                        
                        algorithm = 'port')
            
            mod
          
          }

fit_spline_mod <- function(df){

  mod <- gcFitSpline(time = df$time, data = df$absorbance, gcID = 'best spline', control = grofit.control(interactive = F))

  mod

  }

  df_mod<- if(input$model == 'gomp'){
    
    df_nest %>% mutate(fit_result = map(data, safely(fit_gompertz_mod)))
    
  }else{
    if(input$model == 'biGomp'){
      
        df_nest %>% mutate(fit_result = map(data, safely(fit_biGompertz_mod)))

    } else{
        
      df_nest %>% mutate(fit_result = map(data, safely(fit_spline_mod)))

    }
  }

        # fit data using the above function ...do this safely in case of errors
       df_models <- df_mod %>% 
            ungroup() %>%
            mutate(fit_errors = map(fit_result, 'error')) %>%
             mutate(models = map(fit_result,'result')) %>%
             mutate(fit_error_occured = map(.x = models, .f = is.null), fit_error_occured = unlist(fit_error_occured))
        
        return(df_models)
        #return(df_mod)
        
    })
```


```{r}
    ##### generate the predictions from the models
    df_preds <- reactive({
        
        df_preds <- if(input$model == 'gomp' | input$model == 'biGomp'){
          
          df_models() %>% 
            filter(fit_error_occured == FALSE) %>% 
            mutate(preds = map2(data, models, add_predictions)) %>% 
            unnest(preds)
          
        } else{
          
          tidy_best_spline_preds <- function(df){
            
            tibble(time = df$fit.time,
                   pred = df$fit.data,
                   absorbance = df$raw.data)
            }
          
          df_models() %>% 
            filter(fit_error_occured == FALSE) %>% 
            mutate(preds = map(models, tidy_best_spline_preds)) %>% 
            unnest(preds)
        }
            
        return(df_preds)
    })
```

```{r}

########################################
# GENERATE PARAM ESTIMATES FROM MODELS #
########################################

df_ests <- reactive({
  
  ########################################
  # Getting param estimates depends on model #
  # spline have no conf.int and require parsing grofit output#
  ########################################
  
  df_ests <- if(input$model == 'gomp' | input$model == 'biGomp'){
    
    df_models() %>% 
      filter(fit_error_occured == FALSE) %>% 
      mutate(est_result = map(models, safely(tidy), conf.int = T)) %>% 
      mutate(est_errors = map(est_result, 'error')) %>% 
      mutate(ests = map(est_result,'result')) %>% 
      mutate(est_error_occured = map(.x = ests, .f = is.null), est_error_occured = unlist(est_error_occured)) %>% 
      mutate(ests = ifelse(est_error_occured == T, map(models, tidy, conf.int = F), ests)) %>% 
      unnest(ests)
    
      } else{
        
        tidy_best_spline_ests <- function(df){
          
          tibble(Lag = df$parameters$lambda, Mu = df$parameters$mu, A = df$parameters$A) %>%
            pivot_longer(cols = c(Lag, Mu, A),names_to = 'term',values_to = 'estimate') %>% 
            filter(!is.na(estimate))
          
          }
        
        df_models() %>% 
          filter(fit_error_occured == FALSE) %>% 
          mutate(ests = map(models, tidy_best_spline_ests )) %>% 
          mutate(est_error_occured = T) %>% 
          unnest(ests)

        }
        
  ########################################
  # Convert absolute growth rate Mu_A to relative growth rate Mu_R and doubling time T_d #
  # Mu_A params are different for each model, requiring different if statements #
  ########################################
  
  df_ests <- df_ests %>% mutate(mu_type = ifelse(term %in% c('Mu','Mu_1','Mu_2'), 'Mu_a', 'other'))
  
  if(input$model == 'gomp'){
    
    mu_ests <- df_ests %>% select(well, term, estimate, conf.low, conf.high) %>% filter(term %in% c('A','Mu')) %>% 
      pivot_wider(id_cols = well, names_from = term, values_from = c(estimate, conf.low, conf.high)) %>% 
      mutate(Mu_r      = (estimate_Mu * exp(1)) / estimate_A, 
             Mu_r_high = (conf.high_Mu * exp(1)) / estimate_A, 
             Mu_r_low  = (conf.low_Mu*exp(1)) / estimate_A 
             ) %>% 
      mutate(T_d = log(2) / Mu_r,
             T_d_high = log(2) / Mu_r_high,
             T_d_low = log(2) / Mu_r_low)
          
        mu_r_ests <- mu_ests %>% select(well, estimate = Mu_r, conf.low = Mu_r_low, conf.high = Mu_r_high) %>% mutate(term = "Mu_r", mu_type = 'Mu_r')
        
        t_d_ests <- mu_ests %>% select(well, estimate = T_d, conf.low = T_d_low, conf.high = T_d_high) %>% mutate(term = "T_d", mu_type = 'T_d')
        
        df_ests <- bind_rows(df_ests, mu_r_ests, t_d_ests)
        
  }
  
  if(input$model == 'biGomp'){
    
    mu_ests <- df_ests %>% select(well, term, estimate, conf.low, conf.high) %>% filter(term %in% c('A_1','A_2','Mu_1','Mu_2')) %>% 
      pivot_wider(id_cols = well, names_from = term, values_from = c(estimate, conf.low, conf.high)) %>% 
      mutate(Mu_r_1      = (estimate_Mu_1 * exp(1)) / estimate_A_1, 
             Mu_r_1_high = (conf.high_Mu_1 * exp(1)) / estimate_A_1, 
             Mu_r_1_low  = (conf.low_Mu_1*exp(1)) / estimate_A_1,
             Mu_r_2      = (estimate_Mu_2 * exp(1)) / estimate_A_2, 
             Mu_r_2_high = (conf.high_Mu_2 * exp(1)) / estimate_A_2, 
             Mu_r_2_low  = (conf.low_Mu_2*exp(1)) / estimate_A_2 
             ) %>% 
      mutate(T_d_1 = log(2) / Mu_r_1,
             T_d_1_high = log(2) / Mu_r_1_high,
             T_d_1_low = log(2) / Mu_r_1_low,
             T_d_2 = log(2) / Mu_r_2,
             T_d_2_high = log(2) / Mu_r_2_high,
             T_d_2_low = log(2) / Mu_r_2_low)
          
        mu_r_1_ests <- mu_ests %>% select(well, estimate = Mu_r_1, conf.low = Mu_r_1_low, conf.high = Mu_r_1_high) %>% mutate(term = "Mu_r_1", mu_type = 'Mu_r')
        mu_r_2_ests <- mu_ests %>% select(well, estimate = Mu_r_2, conf.low = Mu_r_2_low, conf.high = Mu_r_2_high) %>% mutate(term = "Mu_r_2", mu_type = 'Mu_r')
        
        t_d_1_ests <- mu_ests %>% select(well, estimate = T_d_1, conf.low = T_d_1_low, conf.high = T_d_1_high) %>% mutate(term = "T_d_1", mu_type = 'T_d')
        t_d_2_ests <- mu_ests %>% select(well, estimate = T_d_2, conf.low = T_d_2_low, conf.high = T_d_2_high) %>% mutate(term = "T_d_2", mu_type = 'T_d')
        
        df_ests <- bind_rows(df_ests, mu_r_1_ests,mu_r_2_ests, t_d_1_ests, t_d_2_ests)
        
    }
  
    if(input$model == 'spline'){
    
      #print(df_ests)
      
      mu_ests <- df_ests %>% select(well, term, estimate) %>% filter(term %in% c('A','Mu')) %>% 
        pivot_wider(id_cols = well, names_from = term, values_from = estimate) %>% 
        mutate(Mu_r = (Mu * exp(1)) / A) %>% 
        mutate(T_d = log(2) / Mu_r)
          
      mu_r_ests <- mu_ests %>% select(well, estimate = Mu_r) %>% mutate(term = "Mu_r", mu_type = 'Mu_r')
          
      t_d_ests <- mu_ests %>% select(well, estimate = T_d) %>% mutate(term = "T_d", mu_type = 'T_d')
      
      df_ests <- bind_rows(df_ests, mu_r_ests, t_d_ests)
        
    }
  
      
      
    
    #print(df_ests)
        
        #add back into df_ests
    

        
    #print(df_ests %>% select(well, term, estimate, mu_type) %>% tail())
    
    return(df_ests)
    
  })


```

```{r}
df_pred_ests <- reactive({
  
  gen_gompertz <- function(a, mu, lag, c, time){
    
    a*exp(-exp(((mu*exp(1))/a)*(lag - time) + 1 )) + c
    
  }
  
  if(input$model == 'gomp'){
    
    #df_pred_ests <- df_preds()
    
    df_ests_low <- df_ests() %>%
      pivot_wider(id_cols = 'well', names_from = 'term', names_prefix = 'low_', values_from = 'conf.low')# %>% 
      #pivot_wider(id_cols = 'well', names_from = 'term', names_prefix = 'high_', values_from = 'conf.high' )
    
    df_ests_high <- df_ests() %>%
      pivot_wider(id_cols = 'well', names_from = 'term', names_prefix = 'high_', values_from = 'conf.high')
    
    df_ests_wide <- left_join(df_ests_low, df_ests_high, by = 'well')
    
    #print(df_ests_wide)
    
    df_pred_ests <- left_join(df_preds(), df_ests_wide, by = 'well') %>%
      mutate(pred_low = gen_gompertz(a = low_A, mu = low_Mu, lag = high_Lag, c = low_C, time = time)) %>% 
      mutate(pred_high = gen_gompertz(a = high_A, mu = high_Mu, lag = low_Lag, c = high_C, time = time))

  }

  if(input$model == 'biGomp'){
    
    #df_pred_ests <- df_preds()
    
    df_ests_wide <- df_ests() %>%
      pivot_wider(id_cols = 'well', names_from = 'term', values_from = 'estimate' )

    df_pred_ests <- left_join(df_preds(), df_ests_wide, by = 'well') %>%
      mutate(pred_1 = gen_gompertz(a = A_1, mu = Mu_1, lag = Lag_1, c = C, time = time)) %>%
      mutate(pred_2 = gen_gompertz(a = A_2, mu = Mu_2, lag = Lag_2, c = C, time = time))

  }
  


    return(df_pred_ests)
})

#renderDataTable(df_pred_ests() %>% select(well, fit_error_occured, A_2))

```



## Plot overviews{.tabset}

### Fit Plots
```{r}

plotOutput("plot")

output$plot <- renderPlot({
  
  
if(input$model == 'gomp'){
  # p <- ggplot(data = df_preds(), aes_string(x=df_preds()$time, y = df_preds()$absorbance)) + 
  #   geom_point(shape = 21) + 
  #   geom_path(aes(y = df_preds()$pred), size = 1, color = 'blue')+ 
  #   geom_path(data = df_pred_ests(), aes(x = df_pred_ests()$time, y = df_pred_ests()$pred_low), size = 1, color = 'red')+
  #   geom_path(data = df_pred_ests(), aes(x = df_pred_ests()$time, y = df_pred_ests()$pred_high), size = 1, color = 'green')+
  #   facet_wrap(~well, scales = 'free') +
  #   labs(x = 'Time (Hrs)', y = 'Absorbance') + theme_bw()
  
  p <- ggplot(data = df_preds(), aes_string(x=df_preds()$time, y = df_preds()$absorbance)) +
    geom_ribbon(data = df_pred_ests(), aes(x = df_pred_ests()$time, ymin = df_pred_ests()$pred_low, ymax = df_pred_ests()$pred_high), fill = 'light gray')+
    geom_point(shape = 21) +
    geom_path(aes(y = df_preds()$pred), size = 1, color = 'blue')+
    facet_wrap(~well, scales = 'free') +
    labs(x = 'Time (Hrs)', y = 'Absorbance') + theme_bw()
  
}
  
  if(input$model == 'biGomp'){

    p <- ggplot(df_pred_ests(), aes(x = df_pred_ests()$time)) +
      geom_point(aes(y = df_pred_ests()$absorbance), shape = 21)+
      geom_path(aes(y = df_pred_ests()$pred), color = 'blue', size = 1) +
      geom_path(aes(y = df_pred_ests()$pred_1), color = 'red', size = 1) +
      geom_path(aes(y = df_pred_ests()$pred_2), color = 'red', linetype = 2, size = 1) +
      facet_wrap(~well, scales = 'free')+
      labs(x = 'Time (Hrs)', y = 'Absorbance') + theme_bw()

  }
  
  if(input$model == 'spline'){
      
      df_ests_spline <- df_ests() %>% pivot_wider(id_cols = well, names_from = term, values_from = estimate)
      
      p <- ggplot(data = df_preds(), aes(x=time, y = pred)) + 
        geom_point(aes(y = absorbance), shape = 21) +
        geom_path(color = 'blue', size = 1)  + 
        geom_abline(data = df_ests_spline, aes(slope = Mu, intercept = -Lag * Mu), color = 'red') + 
        geom_hline(data = df_ests_spline, aes(yintercept = A), color = 'red') + 
        facet_wrap(~well, scales = 'free') + 
        labs(x = 'Time (Hrs)', y = 'Absorbance') + theme_bw()

  }
  
  
  print(p)
        
    })

```

### Error Plots

```{r}

renderPlot({
  
  df_errors <- df_models() %>% filter(fit_error_occured == T)
  
  df_error_plot <- left_join(df_errors, selectedData(), by = 'well')
  
  p <- ggplot(data = df_error_plot, aes_string(x=df_error_plot$time, y = df_error_plot$absorbance)) + 
    geom_point(shape = 21) + 
    facet_wrap(~well, scales = 'free') +
    labs(x = 'Time (Hrs)', y = 'Absorbance') + theme_bw()
  
  print(p)
})

```


## Table{data-height='75%'}

### Error Table

```{r}
renderTable({df_models() %>% select(well, fit_error_occured)})
```

# 4. Download Parameter Estimates {data-orientation=rows}

## {.sidebar data-width='300%'}

### 4. Download Model Results

See instructions for details.

#### Select growth rate units

```{r}

selectInput("mu", NULL, c("Absolute Growth Rate (Mu_a)" = "Mu_a", "Relative Growth Rate (Mu_r)"="Mu_r", "Doubling Time (T_d)" = "T_d"), selected = 'Mu_a')

```

#### Downloads

Parameter estimates of A, Lag, and Mu. Confidence intervals are included for gompertz models.

```{r}
df_ests_output <- reactive({
  
  if(input$model == 'gomp' | input$model == 'biGomp'){
  
  df_ests_output <- df_ests() %>% select(-c(data, fit_result, fit_errors, models, est_result, est_errors))
  
  } else{
    
    if(input$model == 'spline'){
      df_ests_output <- df_ests() %>% select(well, term, estimate)
    }
    
  }
  
  return(df_ests_output)
    
  })

 downloadHandler(
         filename = 'growth_curve_parameter_estimates.csv',
         outputArgs = list(label = 'Parameter estimates'),
         content = function(file) {
             write_csv(df_ests_output(), file)
         }
     )

```

-----

Best fit curves along with raw data.

```{r}

pred_file <- reactive({  
  
  pred_file <- df_preds()

  # if(input$biGomp){
  # 
  #   pred_file <- df_pred_ests()
  #   
  #   }
  
  })


downloadHandler(
  filename = 'growth_curve_model_fits.csv',
  outputArgs = list(label = 'Best fit curves'),
  content = function(file) { write_csv(pred_file() %>% select(-c(data, fit_result, fit_errors, models)), file)}
    )



```

## {data-height='100%'}

### Plot estimates 

```{r}
renderPlot({
        
  #mu_type == input$mu | 
  print(input$mu)
  
  print(df_ests() %>% select(well, term, estimate, mu_type) %>% filter(mu_type == input$mu | mu_type == 'other'), n=50)
  #print(df_ests() %>% select(well, term, estimate, mu_type) %>% filter(mu_type == input$mu | mu_type == 'other') %>% tail())
  
  plot_ests_df <- df_ests() %>% filter(mu_type == input$mu | mu_type == 'other')
  
  p2 <- ggplot(data = plot_ests_df, aes(x=well, y = estimate)) + geom_point(size = 3)+
      facet_wrap(~term, scales = 'free') +
      labs(x = 'Well', y = 'Parameter Estimate') + theme_bw()
  
  if(df_ests()$est_error_occured == F){
    p2 = p2 + geom_pointrange(aes(ymin = conf.low, ymax = conf.high )) 
  }
  
  print(p2)
        
  })

```

## {data-height='25%'}

### Table estimates 

```{r}
renderTable({df_ests() %>% select(well, term, estimate, est_error_occured)})
```

# Parameter Explorer

```{r, echo=FALSE, results='asis'}
#res <- knitr::knit_child('parameter_explorer.Rmd', quiet = TRUE)
#cat(res, sep = '\n')
```


## {.sidebar data-width='300%'}

### Parameters

#### Select growth rate units
```{r}
  
selectInput("mu_2", 
            "", 
            c("Absolute Growth Rate (Mu_a)" = "Mu_a", "Relative Growth Rate (Mu_r)"="Mu_r", "Doubling Time (T_d)" = "T_d"), 
            selected = 'Mu_a')



```

```{r}
  
  conditionalPanel(condition = "input.mu_2 == 'Mu_a'", 
                   numericInput("gen_Mu", "Mu - Absolute growth rate", value = 0.2, min = 0, step = 0.05)
                 )
  
  conditionalPanel(condition = "input.mu_2 == 'Mu_r'", 
                   numericInput("gen_Mu_r", "Mu_R - Relative growth rate", value = 0.2, min = 0, step = 0.05)
                 )
  
  conditionalPanel(condition = "input.mu_2 == 'T_d'", 
                   numericInput("gen_T_d", "T_D - Doubling time", value = 1, min = 0, step = 0.1)
                 )
  

  numericInput("gen_A", "A - Upper asymptote", value = 2, min = 0, step = 0.5)
  numericInput("gen_Lag", "Lag time", value = 4, step = 1)
  numericInput("gen_C", "C - Lower asymptote", 0, step = 0.1)
  numericInput("t_max","Time series length",48, step = 1)
  numericInput("t_step",'Time series step',0.5, step = 0.1)
  checkboxInput("show_data", "Overlay selected user data", F)
  
```


## Plots

### Plots

```{r}
renderPlot({
  
  gen_gompertz <- function(Mu, A, Lag, C = 0, sd = 0, tmax = 48,tstep = 0.5){
  
  t = seq(0,tmax,tstep)  
    
  vals <- A*exp(-exp(((Mu*exp(1))/A)*(Lag - t) + 1)) + C
  
  tibble(t = t, y = vals)
  }
  
  mu_temp <- input$gen_Mu
  
  if(input$mu_2 == 'Mu_r'){mu_temp <- (input$gen_Mu_r*input$gen_A) / exp(1)}
  
  if(input$mu_2 == 'T_d'){mu_temp <- ((log(2)/input$gen_T_d)*input$gen_A) / exp(1)}
  
  #print(df())
  
  df_gen = gen_gompertz(Mu = mu_temp, A = input$gen_A, Lag = input$gen_Lag, C = input$gen_C, tmax = input$t_max, tstep = input$t_step)

  p <- ggplot(df_gen, aes(x = t, y=y)) + geom_path(color = 'blue', size = 1)
  
  if(input$show_data & !is.null(input$file1)){
    
    p <- p+geom_path(data = selectedData(), aes(time, absorbance, group = well))
    
  }
  
  print(p)

})
```